{\bf Practical Concurrency Testing} \\ \normalsize \vspace{1em} \begin{tabular}{c} {\em or: How I Learned to Stop Worrying and Love the Exponential Explosion} \end{tabular}

### 2018-12-06 thesis/education.tex
summary for educ

\chapter{Education}
\label{chap:education}

@@

Concurrency is taught in as many different ways as there are
systems programming classes at universities which teach the subject.
Yet one thing they all have in common is presenting the concurrency bug
as some elusive menace,
against which humanity's best weapon is mere random stress testing.
This chapter will prove stateless model checking's mettle as a better alternative in the educational theatre.

@@

Human subjects research is inherently messy.
Each of the previous section's approaches to evaluating Landslide's educational value
was accompanied by some drawback which prevented it from being perfectly objective science,
but many of them presented tentatively positive results nonetheless,
\revisionminor{and relatively very little negative feedback such as false positives}.
Landslide helped many students find and fix many bugs (\cref{sec:education-eval-bugfinding}),
but making a direct comparison to stress testing, the prior state of the art, is not straightforward.
%and not statistically significant,
Immediate improvement in students' project grades was observed
(\cref{sec:education-eval-grades}),
although statistical significance was lost when attempting to account for selection bias;
and impact on grades alone is a very narrow measure of pedagogical value anyway.
Landslide's debugging power was also found to be statistically significant
for the {\tt thr\_exit\_join} bug in particular.
Students responded overwhelmingly positively in the survey (\cref{sec:education-eval-survey}),
although it is easy to imagine students being equally happy with a ``debugging tool'' that just tells them all the answers.

Nevertheless, I believe each of these partial results taken together
paint an overall picture of success:
students fixed their own bugs {\em and} were happy about it;
students were able to ask for help rather than be deterred by inevitable technical difficulties as far as I know;
students provided intelligent feedback suggesting they truly understood the debugging process;
% well not the student who said "traces should point to where the bug is" but eh can't win em all c.c
et cetera.
%
The rest of this section will discuss the study's limitations and provide some perspective for the future.

@@

The survey was introduced into the study only in time for two semesters' worth of student responses,
after several iterations of collecting only usage and bug report snapshots.
Apart from the obvious improvement of having been surveying students from the beginning,
the following questions could have improved the survey.
\begin{enumerate}
	\item {\em Did you have any technical difficulties with Landslide that I had to intervene on,
		whether in person or over email?}
		(Some students reported this in the ``false positives'' question,
		although fewer than I helped overall, so others must have not mentioned it.)
		Comparing answers to this question across subsequent semesters would give a sense
		of how much Landslide's stability was improving over time
		and whether it was mature enough for unsupervised use in the future.
	\item {\em How well do you feel you understood the research challenges explained in the lecture?}
		and,
		{\em How well do you feel a user should need to understand same in order to benefit from Landslide's bug reports?}
		(Answers on a scale from ``Not at all; Landslide is a magic black box to me''
		to ``I'm ready to work on research in this field myself''.)
		These questions would help fine-tune the lecture material and user guide to maximize student comfort,
		and potentially also corroborate the claim that Landslide is accessible even to novice users.
	\item {\em What additional debugging information would you want displayed on the preemption traces?}
		Knowing now that interpreting preemption traces was a sticking point for many users,
		I would hope to identify the most wished-for features to know what to prioritize improving.
		This could also assess students' understanding of what kinds of information
		would or would not be reasonable for Landslide to record and report.
	\item {\em For each bug Landslide found in your code, how trivial or severe do you feel it was?}
		This would help get a sense of how the students regarded Landslide
		on a spectrum between annoying style checker \revisionminor{and} life-saver,
		and potentially suggest options to make Landslide suppress certain types of bug reports.
		For example, it currently reports spin-wait loops in {\tt mutex\_lock()} as bugs
		with a special message referring students to relevant lecture slides,
		but it's possible refusing to test any code beyond until that bug was fixed
		might have made Landslide less useful overall.
	\item \revision{{\em In addition to finding bugs, did you manage to fully verify your code under
		any of the tests by letting Landslide complete all state spaces before reaching the specified time out?}
		This would provide a sense of how thoroughly students understood the underlying research technique,
		and serve as a follow-up to the ``how many bug fixes verified'' question
		(where students often stopped Landslide midway through after a little while).}
\end{enumerate}

@@

\subsection{Future educational use}

@@

>>>>

\chapter{Education}
\label{chap:education}
% TODO: look at this spacing hack again on fresh eyes and see if it rly improves matters or not
\vspace{1em}

@@

% fix spacing between paragraphs, since sect 5.1 header starts next page
\vspace{3em}
Concurrency is taught in as many different ways as there are
systems programming classes at universities which teach the subject.
Yet one thing they all have in common is presenting the concurrency bug
as some elusive menace,
against which humanity's best weapon is mere random stress testing.
This chapter will prove stateless model checking's mettle as a better alternative in the educational theatre.

@@

This section will discuss the study's limitations and offer some perspective for the future.

@@

The survey was introduced into the study only in time for two semesters' worth of student responses,
after several iterations of collecting only usage and bug report snapshots.
Apart from the obvious improvement of having been surveying students from the beginning,
the following questions could have improved the survey.
\begin{enumerate}
	\item {\em Did you have any technical difficulties with Landslide that I had to intervene on,
		whether in person or over email?}
		(Some students reported this in the ``false positives'' question,
		although fewer than I helped overall, so others must have not mentioned it.)
		Comparing answers to this question across subsequent semesters would give a sense
		of how much Landslide's stability was improving over time
		and whether it was mature enough for unsupervised use in the future.
	\item {\em How well do you feel you understood the research challenges explained in the lecture?}
		and,
		{\em How well do you feel a user should need to understand same in order to benefit from Landslide's bug reports?}
		(Answers on a scale from ``Not at all; Landslide is a magic black box to me''
		to ``I'm ready to work on research in this field myself''.)
		These questions would help fine-tune the lecture material and user guide to maximize student comfort,
		and potentially also corroborate the claim that Landslide is accessible even to novice users.
	\item {\em What additional debugging information would you want displayed on the preemption traces?}
		Knowing now that interpreting preemption traces was a sticking point for many users,
		I would hope to identify the most wished-for features to know what to prioritize improving.
		This could also assess students' understanding of what kinds of information
		would or would not be reasonable for Landslide to record and report.
	\item {\em For each bug Landslide found in your code, how trivial or severe do you feel it was?}
		This would help get a sense of how the students regarded Landslide
		on a spectrum between annoying style checker \revisionminor{and} life-saver,
		and potentially suggest options to make Landslide suppress certain types of bug reports.
		For example, it currently reports spin-wait loops in {\tt mutex\_lock()} as bugs
		with a special message referring students to relevant lecture slides,
		but it's possible refusing to test any code beyond until that bug was fixed
		might have made Landslide less useful overall.
	\item \revision{{\em In addition to finding bugs, did you manage to fully verify your code under
		any tests by letting Landslide complete all state spaces before reaching the specified time out?}
		This would measure how thoroughly students understood the underlying research technique,
		and serve as a follow-up to the ``how many bug fixes verified'' question
		(where students often stopped Landslide midway through after a little while).}
\end{enumerate}

@@

\subsection{Future educational use}
\label{sec:education-future-use}

@@

\section{Summary}

% report on participation %s

\revision{This chapter has introduced Landslide as an educational tool
in the undergraduate operating systems classes at CMU, U. Chicago, and PSU.
I have made the following contributions:

\begin{itemize}
	\item A lecture, polished over seven semesters,
		for introducing stateless model checking as a concurrency testing approach
		in high-level undergraduate classes
		(with many slides admittedly specific to Landslide as they may be)
		(\cref{sec:education-pebbles-recruiting}).
	\item A system for automatically instrumenting student thread library (CMU, PSU) and kernel (U. Chicago)
		project implementations to be compatible with stateless model checking
		(\cref{sec:education-pebbles-instrumentation},
		\cref{sec:education-pintos-instrumentation}).
	\item An experimental procedure for evaluating debugging and verification tools
		as student-accessible teaching aids,
		including a survey for measuring the more intangible human experience
		(\cref{sec:education-survey-pebbles},
		\cref{sec:education-survey-pintos}).
	\item A large evaluation spanning 9 semesters of operating systems classes
		(\cref{sec:education-eval-bugfinding}),
		2 of which were held at other universities besides CMU,
		in which 145 students used Landslide,
		a majority of which found bugs (\cref{sec:education-eval-bugfinding}),
		and a vast majority of which expressed via the survey that they had a good experience
		(\cref{sec:education-eval-survey}).
	\item A thorough statistical analysis of Landslide's effect on student submission quality
		(\cref{sec:education-eval-grades}),
		evaluating both impact on ultimate project grades
		and student tendency to submit any of four well-known case study bugs.
		Ultimately, while positive trends were observed,
		this study concluded that a statistically-significant conclusion could not be drawn
		while simultaneously compensating for several obvious sources of possible bias.
		Nevertheless, this level of statistical rigor is not yet widespread in computer science research,
		so I would hope this study may set a precedent that help mature the field.
	\item Concrete action taken to establish Landslide
		as a permanent fixture of 15-410's grading process at CMU,
		student-oriented user guides should the course staff wish it to remain available as a debugging tool,
		as well as general recommendations for any Pintos-based OS class
		at other universities to use it for grading as well
		(\cref{sec:education-future-use}).
\end{itemize}
}

Human subjects research is inherently messy.
Each %of the previous section's
\revisionminor{individual}
approaches to evaluating Landslide's educational value
was accompanied by some drawback which prevented it from being perfectly objective science,
but many of them presented tentatively positive results nonetheless,
\revisionminor{and relatively very little negative feedback such as false positives}.
Landslide helped many students find and fix many bugs (\cref{sec:education-eval-bugfinding}),
but making a direct comparison to stress testing, the prior state of the art, is not straightforward.
%and not statistically significant,
Immediate improvement in students' project grades was observed
(\cref{sec:education-eval-grades}),
although statistical significance was lost when attempting to account for selection bias;
and impact on grades alone is a very narrow measure of pedagogical value anyway.
Landslide's debugging power was also found to be statistically significant
for the {\tt thr\_exit\_join} bug in particular.
Students responded overwhelmingly positively in the survey (\cref{sec:education-eval-survey}),
although it is easy to imagine students being equally happy with a ``debugging tool'' that just tells them all the answers.
%
Nevertheless, I believe each of these partial results taken together
paint an overall picture of success:
students fixed their own bugs {\em and} were happy about it,
students were able to ask for help rather than be deterred by inevitable technical difficulties, %as far as I know;
and students provided intelligent feedback suggesting they truly understood the debugging process.
% well not the student who said "traces should point to where the bug is" but eh can't win em all c.c
%et cetera.
